---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}

knitr::opts_chunk$set( #set default chunk options to not show the R code
  echo = FALSE
)
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

```{r}
smaller %>%
  ggplot(aes(x = carat, color = color)) + 
  geom_freqpoly(binwidth = 0.01)
```

```{r}
smaller %>%
  ggplot(aes(x = carat, color = cut)) + 
  geom_freqpoly(binwidth = 0.01)
```

***

27.3.1 Exercise 3

***

***

27.4.7 Exercise 1

***

***

# Nava Nizard

---

## Education
* CUNY Hunter College (**2024 - 2028**)
  * Bachelor's of Arts in Computer Science

---

## Employment
* Microsoft: Data Science Research Fellow [^1] (June **2026**)
  * Learnt R and the basics of statistics and data science.

* InterDigital Inc.: Software Engineer Intern (Jan **2026**)
  * Developed SQL scripts for a CLM migration.
  * Optimized the scripts for fast execution time.

---

## Quote
> You miss 100% of the shots you don't take

[^1]: This is techinically not an employment in the regular sense of the word.

***
