---
title: "Testing Regression Model for Citibike Trips"
author: "Nava Nizard (Hunter College)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
repo = "https://cran.r-project.org"

install.packages("httpgd", repos = repo)
library(tidyverse)
library(modelr)
library(scales)
library(httpgd)
httpgd::hgd()

theme_set(theme_bw())

knitr::opts_chunk$set(echo = TRUE)

setwd("c:/Users/nava/Documents/coursework/week4/")

load("c:/Users/nava/Documents/coursework/week3/final_model.RData")
```

## Load data

```{r}
trips_per_day <- read_tsv("trips_per_day_2015.tsv")
weather <- read_csv("weather_2015.csv") %>% 
    rename("ymd" = "DATE", 
           "prcp" = "PRCP", 
           "snow" = "SNOW",
           "snwd" = "SNWD", 
           "tmax" = "TMAX", 
           "tmin" = "TMIN")
trips_with_weather <- left_join(trips_per_day, weather, by = "ymd") %>%
    select(-STATION, -NAME, -TAVG)
```

## Feature Engineer New Data Set
```{r feature-engineer}
# make sure to have same fields used in the model
trips_with_weather <- trips_with_weather %>% filter(year(ymd) <= 2015)
trips_with_weather <- trips_with_weather %>%
    mutate(tmin = tmin * 0.1, tmax = tmax * 0.1) %>%
    mutate(season = case_when(month(ymd) == 12 | month(ymd) == 1 | month(ymd) == 2 ~ "winter",
                              month(ymd) == 3  | month(ymd) == 4 | month(ymd) == 5 ~ "spring",
                              month(ymd) == 6  | month(ymd) == 7 | month(ymd) == 8 ~ "summer",
                              month(ymd) == 9  | month(ymd) == 10 | month(ymd) == 11 ~ "fall")) %>%
    mutate(day_of_the_week = wday(ymd, label = TRUE)) %>%
    group_by(season) %>%
    mutate(rain_lvl = case_when(prcp <= mean(prcp) ~ "low",
                            prcp > mean(prcp) ~ "high")) %>%
    mutate(avg_temp = (tmin + tmax) / 2) %>%
    mutate(avg_temp = case_when(avg_temp <= mean(avg_temp) ~ "cold",
                                avg_temp > mean(avg_temp) ~ "hot"))

# View(trips_with_weather)
# View(train)       
```

```{r testing}
test_predictions <- predict(final_model, trips_with_weather) 
test_rmse <- rmse(final_model, trips_with_weather) #8813 trips
rsquare(final_model, trips_with_weather) #70%
```

```{r plot-data}
plot_test_data <- data.frame(ymd = trips_with_weather$ymd, actual_trips = trips_with_weather$num_trips, test_predictions)

ggplot(plot_test_data, aes(x = ymd)) +
    geom_line(aes(y = test_predictions, color = "Predicted")) +
    geom_point(aes(y = actual_trips, color = "Actual")) + 
    scale_color_manual(values = c("Predicted" = "black",
                                   "Actual" = "blue")) +
    labs(x = 'Date', 
         y = 'Number of Trips',
         title = 'Predicted VS Actual Number of Trips',
         color = "",)

ggplot(plot_test_data, aes(x = test_predictions, y = actual_trips)) +
    geom_point() +
    labs(x = 'Predicted Number of Trips', 
         y = 'Actual Number of Trips',
         title = 'Model Accuracy in Predicting Number of Trips')
```

## Conclusion
When I first tested my model on the test set, I got a very large RMSE of about 120,000 trips 
and noticed some huge spikes in the predictions. After investigating, I realized that the issue 
was coming from my feature engineering. I had originally created `avg_temp` and `rain_lvl` using 
averages across the entire year instead of grouping by season first. This caused some days to be 
classified in a way that created extra fluctuations in the predictions.

After fixing this by grouping by season before calculating these variables, the RMSE decreased 
a lot to `r rmse(final_model, trips_with_weather)` trips. (I know that you're not supposed to change your model after looking 
at the test results, but in this case I was correcting a mistake in my feature engineering rather 
than intentionally tuning the model to perform better on the test set.)

There are still some large fluctuations during the summer months, which suggests that there 
may be other factors affecting Citi Bike usage that are not included in my model. Overall, 
the model improved a lot after fixing the feature engineering issue, but there is still room 
for improvement.