---
title: "Predicting Daily Citibike Trips"
author: "Nava Nizard (Hunter College)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
repo = "https://cran.r-project.org"

install.packages("httpgd", repos = repo)
library(tidyverse)
library(modelr)
library(scales)
library(httpgd)
httpgd::hgd()

theme_set(theme_bw())

knitr::opts_chunk$set(echo = TRUE)

setwd("c:/Users/nava/Documents/coursework/week3/")
```

## Load data

```{r}
trips_per_day_total <- read_tsv("trips_per_day.tsv")
```

## Train/test split
Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
set.seed(42) # makes the random sampling reproducible

# STEP 1: FEATURE ENGINEERING THE DATA:
trips_per_day_total <- trips_per_day_total %>%
    mutate(season = case_when(month(ymd) == 12 | month(ymd) == 1 | month(ymd) == 2 ~ "winter",
                              month(ymd) == 3  | month(ymd) == 4 | month(ymd) == 5 ~ "spring",
                              month(ymd) == 6  | month(ymd) == 7 | month(ymd) == 8 ~ "summer",
                              month(ymd) == 9  | month(ymd) == 10 | month(ymd) == 11 ~ "fall")) %>%
    mutate(rain_lvl = case_when(prcp <= mean(prcp) ~ "low",
                            prcp > mean(prcp) ~ "high")) %>%
    mutate(avg_temp = (tmin + tmax) / 2) %>%
    mutate(avg_temp = case_when(avg_temp <= mean(avg_temp) ~ "cold",
                                avg_temp > mean(avg_temp) ~ "hot")) %>%
    mutate(day_of_the_week = wday(ymd, label = TRUE))
View(trips_per_day_total)

# STEP 2: SPLIT THE DATA [80 | 10 | 10] -> [train | validate | test]
num_of_days <- nrow(trips_per_day_total)
rows <- sample(num_of_days, replace = FALSE) # randomize the row indices
n_train <- floor(num_of_days * 0.8) # number of rows for the training data (80%)
n_validate <- floor(num_of_days * 0.1) # number of rows for the validate data (10%)
train <- trips_per_day_total[rows[1:n_train],] # train = 80% of data (from 0-80)
validate <- trips_per_day_total[rows[(n_train + 1):(n_train + n_validate)],] # validate = 10% of data (from 80-90)
test <- trips_per_day_total[rows[(n_train + n_validate + 1):num_of_days],] # test = 10% of data (from 90-100)
```

## Baseline model: temperature only
Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data

K <- 1:8
train_err <- c() # vector with training RMSE for each degree
validate_err <- c() # vector with validatation RMSE for each degree

for (k in K) {
    tmin_model <- lm(num_trips ~ poly(tmin, k, raw = T), data = train)

    train_err[k] <- sqrt(mean((predict(tmin_model, train) - train$num_trips)^2)) # RMSE = sqrt(mean((predicted_val - actual_val)^2)
    validate_err[k] <- sqrt(mean((predict(tmin_model, validate) - validate$num_trips)^2))
}

# Plot the training and validation error as a function of the polynomial degree
plot_error_data <- data.frame(K, train_err, validate_err) %>%
    pivot_longer(names_to = 'error_type', values_to = 'RMSE', cols = -K)

plot_error_data %>%
    ggplot(aes(x = K, y = RMSE, color = error_type)) +
    geom_line() +
    scale_x_continuous(breaks = K) +
    labs(x = 'Degree of Polynomial', 
         y = 'Root Mean Square Error (RMSE)',
         color = 'Data Type',
         title = 'Training and Validation RMSE by Polynomial Degree') +
    scale_color_discrete(labels = c(train_err = "Training Data",
                                    validate_err = "Validation Data"))
# Conlusion: Choose k = 4
```

## Feature engineering
Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}
# I did the feature engineering in an earlier step, before splitting the data.
```

## Model comparison
Try different models and compare their validation performance.

```{r}
seasons_model <- lm(num_trips ~ season, data = train)
seasons_r_sqr <- summary(seasons_model)$r.squared # 0.57 
rmse(seasons_model, validate) < validate_err[4] # FALSE

rain_lvl_model <- lm(num_trips ~ rain_lvl, data = train)
rain_lvl_r_sqr <- summary(rain_lvl_model)$r.squared # 0.07 
rmse(rain_lvl_model, validate) < validate_err[4] # FALSE

avg_temp_model <- lm(num_trips ~ avg_temp, data = train)
avg_temp_r_sqr <- summary(avg_temp_model)$r.squared # 0.66 GOOD
rmse(avg_temp_model, validate) < validate_err[4] # FALSE

day_of_the_week_model <- lm(num_trips ~ day_of_the_week, data = train)
day_of_the_week_r_sqr <- summary(day_of_the_week_model)$r.squared # 0.05 
rmse(day_of_the_week_model, validate) < validate_err[4] # FALSE

seasons_and_avg_temp_model <- lm(num_trips ~ (season * avg_temp), data = train) # * slightly higher r_sqrt than +. 
seasons_and_avg_temp_r_sqr <- summary(seasons_and_avg_temp_model)$r.squared # 0.71 
rmse(seasons_and_avg_temp_model, validate) < validate_err[4] # FALSE

seasons_and_avg_temp_and_dow_model <- lm(num_trips ~ (season * avg_temp) + day_of_the_week, data = train) # 
seasons_and_avg_temp_and_dow_r_sqr <- summary(seasons_and_avg_temp_and_dow_model)$r.squared # 0.76 
rmse(seasons_and_avg_temp_and_dow_model, validate) < validate_err[4] # FALSE

all_model <- lm(num_trips ~ (season * avg_temp * rain_lvl) + day_of_the_week, data = train)
all_r_sqr <- summary(all_model)$r.squared # 0.82 
rmse(all_model, validate) < validate_err[4] # TRUE

all2_model <- lm(num_trips ~ (season * avg_temp * rain_lvl) + day_of_the_week + snow, data = train)
all2_r_sqr <- summary(all2_model)$r.squared # 0.83 
rmse(all2_model, validate) < validate_err[4] # TRUE

all3_model <- lm(num_trips ~ (season * avg_temp * rain_lvl * tmin) + day_of_the_week + snow, data = train)
all3_r_sqr <- summary(all3_model)$r.squared # 0.8938 
rmse(all3_model, validate) < validate_err[4] # TRUE

all4_model <- lm(num_trips ~ (season * avg_temp * rain_lvl * tmin) + prcp + day_of_the_week + snow, data = train)
all4_r_sqr <- summary(all4_model)$r.squared # 0.8987 
rmse(all4_model, validate) < validate_err[4] # TRUE

all5_model <- lm(num_trips ~ (season * avg_temp * rain_lvl * tmin) + prcp + day_of_the_week + snwd, data = train)
all5_r_sqr <- summary(all5_model)$r.squared # 0.90
rmse(all5_model, validate) < validate_err[4] # TRUE

```

## Final model
Pick your best model and plot the results.

```{r}
final_model <- lm(num_trips ~ (season * avg_temp * rain_lvl * tmin) + prcp + day_of_the_week + snwd, data = train)
final_r_sqr <- summary(final_model)$r.squared # 0.9018
rmse(final_model, validate) < validate_err[4] # TRUE

predicted_trips <- predict(final_model, test) 
plot_test_data <- data.frame(ymd = test$ymd, actual_trips = test$num_trips, predicted_trips)

ggplot(plot_test_data, aes(x = ymd)) +
    geom_line(aes(y = predicted_trips, color = "Predicted")) +
    geom_point(aes(y = actual_trips, color = "Actual")) + 
    scale_color_manual(values = c("Predicted" = "black",
                                   "Actual" = "blue")) +
    labs(x = 'Date', 
         y = 'Number of Trips',
         title = 'Predicted VS Actual Number of Trips',
         color = "",)

ggplot(plot_test_data, aes(x = predicted_trips, y = actual_trips)) +
    geom_point() +
    labs(x = 'Predicted Number of Trips', 
         y = 'Actual Number of Trips',
         title = 'Model Accuracy in Predicting Number of Trips')
```

## Save model

```{r}
 #.RData file using the save function. 
save(final_model, file = "final_model.RData")
```

## Test set evaluation
**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
load("final_model.RData")
test_rmse <- rmse(final_model, test)
test_rmse

test_rmse >= rmse(final_model, validate) # TRUE, which is good!
```

My final model had a test RMSE of `r rmse(final_model, test)` trips per day. This is much lower than the 
baseline model's RMSE of `r validate_err[4]`, so adding more features improved the predictions. The test RMSE was also 
very close to the validation RMSE of `r rmse(final_model, validate)`, which suggests that the model performed 
similarly on new data and was not overfit to the training data.

I think the model will perform about the same on future data, although it could be slightly worse because 
differnet factors may be different from the data used to train the model.
