---
title: "Predicting Daily Citibike Trips"
author: "Nava Nizard (Hunter College)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
repo = "https://cran.r-project.org"

install.packages("httpgd", repos = repo)
library(tidyverse)
library(modelr)
library(scales)
library(httpgd)
httpgd::hgd()

theme_set(theme_bw())

knitr::opts_chunk$set(echo = TRUE)

setwd("c:/Users/nava/Documents/coursework/week3/")
```

## Load data

```{r}
trips_per_day_total <- read_tsv("trips_per_day.tsv") %>%
    select(ymd, num_trips, tmin)
```

## Train/test split
Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
trips_per_day <- (store 90%)
trips_per_day_test_data <- (store other 10%)
```

## Baseline model: temperature only
Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data
```

## Feature engineering
Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}

```

## Model comparison
Try different models and compare their validation performance.

```{r}

```

## Final model
Pick your best model and plot the results.

```{r}

```

## Save model

```{r}

```

## Test set evaluation
**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
```
